---
title: "volleyball.rmd"
output: html_document
date: "2022-09-28"
---

```{r}
library(tidyverse)
```

#The two additional factors I wanted to see from the quiz were blocks and game score. To reiterate, blocks are an essential part of the game, and within the data we were provided, there are several components of blocks — solo blocks, block assists and block errors. We already have kills and aces (two offensive metrics) and digs (a defensive stat), so I felt like adding a second defensive metric in blocks would help show which teams were the strongest. I also wanted to leave in kills, aces and digs because each of those metrics is intergral to the sport. 

#We also have the game score data in team_score and opponent_score. I wanted to include this information because it is one way for us to measure dominance with the data we were given. For example, a team that wins a majority of its games 3-0 is a stronger unit than a team that wins 3-2. Granted, strength of opponent is a factor in this, but seeing as we don't have that data, team and opponent score seem like good ways to see which teams were the most dominant.

```{r setup, include=FALSE}
match_data <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensVolleyballData/main/data/ncaa_womens_volleyball_matchstats_2022.csv")

head(match_data)

teamquality <- match_data %>% 
  select(team, kills, digs, aces, block_solos, block_assists, b_err, team_score, opponent_score)

teamtotals <- teamquality %>% 
  group_by(team) %>% 
  summarise(
    KillsAvg = mean(kills), 
    DigsAvg = mean(digs), 
    AcesAvg = mean(aces),
    BSoloAvg = mean(block_solos),
    BAssAvg = mean(block_assists),
    BErrAvg = mean(b_err),
    TeamScoreAvg = mean(team_score),
    OppScoreAvg = mean(opponent_score),
    ) 

#These stats all seem important
teamzscore <- teamtotals %>% 
  mutate(
    killszscore = as.numeric(scale(KillsAvg, center = TRUE, scale = TRUE)),
    digszscore = as.numeric(scale(DigsAvg, center = TRUE, scale = TRUE)),
    aceszscore = as.numeric(scale(AcesAvg, center = TRUE, scale = TRUE)),
    soloblockszscore = as.numeric(scale(BSoloAvg, center = TRUE, scale = TRUE)),
    blockassistzscore = as.numeric(scale(BAssAvg, center = TRUE, scale = TRUE)),
    blockerrorzscore = as.numeric(scale(BErrAvg, center = TRUE, scale = TRUE)),
    teamscorezscore = as.numeric(scale(TeamScoreAvg, center = TRUE, scale = TRUE)),
    oppscorezscore = as.numeric(scale(OppScoreAvg, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = killszscore + digszscore + aceszscore + soloblockszscore + blockassistzscore + blockerrorzscore + teamscorezscore + oppscorezscore
    ) %>% ungroup() 

teamzscore %>% arrange(desc(TotalZscore))

#There are too many stats involved with blocks!
teamzscore2 <- teamtotals %>% 
  mutate(
    killszscore = as.numeric(scale(KillsAvg, center = TRUE, scale = TRUE)),
    digszscore = as.numeric(scale(DigsAvg, center = TRUE, scale = TRUE)),
    aceszscore = as.numeric(scale(AcesAvg, center = TRUE, scale = TRUE)),
    teamscorezscore = as.numeric(scale(TeamScoreAvg, center = TRUE, scale = TRUE)),
    oppscorezscore = as.numeric(scale(OppScoreAvg, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = killszscore + digszscore + aceszscore + teamscorezscore + oppscorezscore
    ) %>% ungroup() 

teamzscore2 %>% arrange(desc(TotalZscore))

#Third time's the charm?
teamzscore3 <- teamtotals %>% 
  mutate(
    teamscorezscore = as.numeric(scale(TeamScoreAvg, center = TRUE, scale = TRUE)),
    oppscorezscore = as.numeric(scale(OppScoreAvg, center = TRUE, scale = TRUE)) * -1,
    TotalZscore = teamscorezscore + oppscorezscore
    ) %>% ungroup() 

teamzscore3 %>% arrange(desc(TotalZscore))

big_ten <- c("Nebraska Cornhuskers", "Minnesota Golden Gophers", "Iowa Hawkeyes", "Wisconsin Badgers", "Illinois Fighting Illini", "Northwestern Wildcats", "Purdue Boilermakers", "Indiana Hoosiers", "Michigan Wolverines", "Michigan St. Spartans", "Ohio St. Buckeyes", "Penn St. Nittany Lions", "Maryland Terrapins, Terps", "Rutgers Scarlet Knights")

teamzscore %>% 
  filter(team %in% big_ten) %>% 
  arrange(desc(TotalZscore)) %>%
  select(team, TotalZscore, teamscorezscore, oppscorezscore)

teamzscore %>% 
  filter(team == "Maryland Terrapins, Terps") %>% 
  arrange(desc(TotalZscore)) %>%
  select(team, TotalZscore)
```

If we run the teamzscore block of code that contains the 14 Big Ten teams, these are the results we get:
1. Purdue (6.54)
2. Michigan (4.67)
3. Maryland (4.62)
4. Penn St. (3.83)
5. Mich St. (3.04)
6. Wisconsin (3.03)
7. Northwestern (2.01)
8. Nebraska (1.69)
9. Illinois (1.55)
10. Minn (0.96)
11. Rutgers (0.89)
12. Indiana (0.84)
13.Iowa (-1.46)
14. OSU (-3.73)

This seems kind of weird. This model loves Purdue but hates OSU. In adding team and opponent score, this model likes squads with strong records, and Purdue has a bunch of 3-0 wins, whereas OSU has a lot of 0-3 losses combined with 3-0 wins. As such, Purdue has a strong teamscorezscore (1.46) and OSU has a weak one (-0.11). However, OSU is one of the top teams in the country. Might strength of opponent matter? More on that in a bit.

#Who are the highest-performing teams in the NCAA (based on my new-and-improved TotalZScores - this is using the data from teamzscore)? 

Here are the top 10 highest-performing teams in the NCAA with their z-scores:
1. UCF Knights (9.29)
2. UNLV Lady Rebels (9.14)
3. New Mexico Lobos (8.59)
4. Arkansas Razorbacks (8.30)
5. Oregon Ducks (8.21)
6. Wright St. Raiders (7.89)
7. South Dakota Coyotes (7.87)
8. UTEP Miners (7.64)
9. Idaho St. Bengals (7.57)
10. Auburn Tigers (7.49)

And here is the AVCA Coaches ranking of the top 10 teams with their zscores:
1. Texas (2.43)
2. Louisville (5.15)
3. Nebraska (1.69)
4. Univ. of SD (4.24)
5. Purdue (6.54)
6. OSU (-3.73)
7. Minnesota (0.96)
8. Wisconsin (3.03)
9. Stanford (-0.12)
10. Pitt (4.37)

#Clearly, despite my best intentions, the highest-performing teams in the NCAA based on my z-scores aren't the actual best ten teams in the country according to the rankings. Rankings are subjective, but the 10 teams listed in the AVCA are some of the best in the nation. WHY IS THIS...? I have a couple thoughts. 

#First, let's just take the top teams in each list — UTEP and Texas — and compare the seasons they are having just to see why we got the teams we did in the first list. 

```{r}
teamzscore %>% 
  filter(team == "UCF Knights")

#Here are UTEP's numbers in averages:
#Kills: 52.8
#Digs: 51.5
#Aces: 5.4
#BlockSolo: 2.1
#BlockAssist: 15.2
#BlockError: 2
#TeamScore: 3
#OpponentScore: 0.7
  
teamzscore %>% 
  filter(team == "Texas Longhorns") 

#Here are Texas's numbers in averages:
#Kills: 49.6
#Digs: 46.6
#Aces: 5.8
#BlockSolo: 0.7
#BlockAssist: 12.2
#BlockError: 1.1
#TeamScore: 3
#OpponentScore: 0.5
```

Woah, UCF looks pretty darn competitive here! But wait, let's glance at the final two categories. UCF is a perfect 12-0 and is a solid team. And of their wins, eight have been by a score of 3-0. Texas is also perfect, holding an 11-0 record with seven wins coming at 3-0. Texas has also defeated multiple team that are nationally ranked so far. Let's look at these two teams' average team and opponent scores. We know how dominant Texas is, and they win their matches by an average score of 3-0.5, but UCF is pretty close going 3-0.7. while the Miners win by an average score of 2.2-2.1. If UCF is ranked first in our totalzscore list but Texas is ranked number one by the poll, we know there is something to using game scores to evaluate how good a team is. We could also ask how much we dilute a team's talent by adding in all these additional statistics.

Which leads us to...

Another issue with these z-scores is that we're assuming that each one of these metrics is equally related to victories. Is it possible that aces are more important than blocks? Or are kills more valuable than block assists? And maybe having three different categories for blocks hurts some teams and helps others. For example, a team might have one player who is an incredibly dominant blocker so it doesn't get many block assists or block errors, this hurting their zscore even though that team probably wins a lot of points through that one player. Maybe adding in all the nitty gritty parts of blocks messes up our totalzscore.

Another important element to consider is that there is no way to adjust for strength of schedule within our data. Based on a total z-score, UCF might come out ahead of Texas, even though a game between the two teams would most likely end in a Texas victory. Interestsingly enough, UCF defeated Kansas 3-0 (WOAH) earlier this season while Texas narrowly escaped with a 3-2 win. AND both games were played in Kansas. This probably contributes to UCF ranking above Texas, but I do want to mention that just using game score might not be that effective because a team might beat a good oppoennt 3-2 while another team might win 3-0 against a weaker oppoennt and get favored by our model.  

#Where does Maryland rank on the Z-score scale and what teams are similar to it? 
Maryland's z-score is 2.7, which makes it the 79th best team in the country according to these metrics. By these metrics, Maryland is a better overall team than all of the teams that appear in the AVCA list...When we look at the Big Ten standings this season, Maryland is 0-2 with losses to Illinois and Indiana. Last season, the Terps were 7-13 in conference play and finished ninth. 

The nine teams that are most similar to Maryland using these metrics are the Grand Canyon Lopes, Drake Bulldogs, Southern Miss. Golden Eagles, Kent St. Golden Flashes, Georgia St. Panthers, St. John's (NY) Red Storm, Bellarmine Knights, Chicago St. Cougars, Miami (FL) Hurricanes.

#This was my first attempt at this assignment. Here, I tried to follow the steps from the tutorial. I got to the very end but realized I was working with match data as opposed to using data that had been compressed for each team, so I started semi-anew (seen above). Still, I thought I would leave it here to show my first try.
```{r}
perset_stats <- match_data %>% 
  mutate(
    killsperset = kills/s,
    digsperset = digs/s,
    acesperset = aces/s,
    soloblocksperset = block_solos/s,
    blockassistsperset = block_assists/s,
    blockerrorsperset = b_err/s,
    teamscoreperset = team_score/s,
    oppscoreperset = opponent_score/s
    ) %>%
  select(date, team, killsperset, digsperset, acesperset, soloblocksperset, blockassistsperset, blockerrorsperset, teamscoreperset, oppscoreperset)

head(perset_stats)

teamzscore <- perset_stats %>%
  group_by(team) %>% 
  mutate(
    killszscore = as.numeric(scale(killsperset, center = TRUE, scale = TRUE)),
    digszscore = as.numeric(scale(digsperset, center = TRUE, scale = TRUE)),
    aceszscore = as.numeric(scale(acesperset, center = TRUE, scale = TRUE)),
    soloblockszscore = as.numeric(scale(soloblocksperset, center = TRUE, scale = TRUE)),
    blockassistzscore = as.numeric(scale(blockassistsperset, center = TRUE, scale = TRUE)),
    blockerrorzscore = as.numeric(scale(blockerrorsperset, center = TRUE, scale = TRUE)),
    teamscorezscore = as.numeric(scale(teamscoreperset, center = TRUE, scale = TRUE)),
    oppscorezscore = as.numeric(scale(oppscoreperset, center = TRUE, scale = TRUE)),
    TotalZscore = killszscore + aceszscore + digszscore + soloblockszscore + blockassistzscore + blockerrorzscore + teamscorezscore + oppscorezscore
    ) %>% ungroup() 

head(teamzscore)

teamzscore %>% arrange(desc(TotalZscore))

teamzscore %>% 
  filter(team == "Maryland Terrapins, Terps") %>% 
  arrange(desc(TotalZscore)) %>%
  select(date, team, TotalZscore)
```