---
title: "wbb.Rmd"
output: html_document
date: "2022-11-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(usdata)
```

```{r}
wbb_rosters <- read_csv("https://github.com/dwillis/wbb-rosters/raw/main/rosters_2022-23.csv")
```

#Things to clean
-Standardize positions
-Standardize hometowns
-Drop where name is NA
-Standardize year column
-Standardize height (5'10" or 5-10)

#Columns to add
-State
-Country
-Height in inches


```{r}
wbb_rosters %>% distinct(hometown)
```

```{r}
cleaned_wbb_rosters <- wbb_rosters %>%
  mutate(hometown_clean = str_to_upper(hometown))

cleaned_wbb_rosters <- cleaned_wbb_rosters %>%
  separate(col=hometown_clean, into=c("new_hometown", "state"), sep=",", extra="merge")

```

```{r}
cleaned_wbb_rosters %>% distinct(state)
```

```{r}
cleaned_wbb_rosters <- cleaned_wbb_rosters %>% 
  mutate(state = str_trim(state), state = str_replace(state, '\\.','')) %>% 
  mutate(state = case_when(
     is.na(abbr2state(state)) ~ state,
     TRUE ~ str_to_upper(abbr2state(state)))
  )
```


```{r}
cleaned_wbb_rosters %>% group_by(position) %>% summarise(players = n()) %>% arrange(desc(players))

cleaned_wbb_rosters <- cleaned_wbb_rosters %>%
  mutate(position_clean = str_to_upper(position)) %>%
  mutate(position_clean = case_when(position_clean == 'G' ~ 'GUARD',
                                    position_clean == 'F' ~ 'FORWARD',
                                    position_clean == 'G/F' ~ 'GUARD/FORWARD',
                                    position_clean == 'F/C' ~ 'FORWARD/CENTER',
                                    position_clean == 'C/F' ~ 'FORWARD/CENTER',
                                    position_clean == 'C' ~ 'CENTER',
         TRUE ~ position_clean))

positions <- cleaned_wbb_rosters %>% group_by(position, position_clean) %>% summarise(count = n()) %>% arrange(desc(count))

cleaned_wbb_rosters %>% filter(position == "Guard/Post")

write_csv(positions, "~/Downloads/positions.csv")
```

```{r}
positions_cleaned <- read_csv("~/Documents/772/jour479x_fall_2022/data/wbb_positions.csv") %>%
  select(-count)

wbb_rosters <- wbb_rosters %>% inner_join(positions_cleaned, by = "position") 
```

```{r}
wbb_rosters %>% 
  group_by(position_clean) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

```{r}
wbb_rosters21_22 <- read_csv("https://github.com/dwillis/wbb-rosters/raw/main/rosters_2021-22.csv")

wbb_rosters21_22 %>% 
  group_by(position) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
```

What needs to be done to the positions column of this data is as follows:

One of the main differences in this data set is that we have a bunch of players names. Thankfully, each one is listed only once, so they should be easy to remove. 

We also have around 12-15 results for class year and four results for Combo, but again, those shouldn't take too long to remove. 

The only additional entries we have seem to be positions that are thrice listed. For example, PG/SG/SF. We also have entries that are as follows: FWD/C. But we don't have anything that would be hard to distinguish if we went through and cleaned the data.

Those are the main issues with the position information I saw from 21-22 data.

In total, I would say this impacts no more than 30 rows — again, a lot of the funky stuff has just one entry.

In terms of the other roster data — hometown, year, high school, previous school and height — those all need to be standardized, but that's the main concern.