---
title: "background.rmd"
output: html_document
date: "2022-10-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##DATA BACKGROUNDING (10/08/22)

#Here is my original question: At what point should a team decide to "blow up" its roster? 
#I want to expand on this a little bit to ask the question of how long before the Nats — in light of their 107-loss (a Washington baseball record) season — are relevant again. I'll go into more detail below.
A) I want to start by asking the question of when teams decide to blow up their rosters. To do this, I will look at recent sell-offs and what the records of those teams were. I will use the record of when the majority of the "big moves" involving notable players were made. This will help me get an idea of when teams begin the firesale. Most teams make several moves, but the big ones seem to happen in quick succession. This will help show if there's a point when teams have decided to blow up their rosters. By looking at records, I might be able to observe a pattern of when teams throw in the towel and start planning for the future.

B) I want to look at how many years it takes before a team is good again. I will define "good" in two ways: 1. When a team that has had a losing record finally flips to a winning record, at least 82-80, and 2) When a team that has been struggling makes the playoffs. The important metrics to look at here are win percentage and postseason appearances. For the postseason category, I'm not interested in teams, like the Phillies, that have spent big in recent years and made the postseason this year after a lengthy drought. I want to look at teams (Orioles, Royals, Padres, Marlins, Astros, Pirates, Braves, Cubs) that built through trades and the draft to find success and return to a competitive brand of baseball.

C) One of the main components of this investigation will be to look at the farm systems of certain teams to project when they will return to relevancy. While this is impossible to predict, we can look at players at the AAA (Rochester Red Wings) and AA (Harrisburg Senators) level to see who is performing at above-league rates who might contribute at the major league level. 

This is an oversimplified way of judging whether or not a player will be good, but the thought process is as follows: If player X is above-average at the minor league level, then he will contribute at the pro level. I can also look at roughly how many players per team were drafted/acquired via trade who contributed (defined by a player who has played significant games, > 80 and at-bats, > 320) for their teams. The last part of this question is how many contributing players were signed via free agency, and what were the contracts of those players? This portion looks at how important free agency and spending are in a rebuild. This part of my investigation will look at players drafted, players acquired via trade and players signed in the offseason.

#Does that data already exist? Can you get it?
All the data I will be using (records, years to relevancy and player data) is available, and I should be able to get it without having to spend any money or submit any requests.

#Do you need assistance in getting it?
I'm not sure how I could get the data for the first two queries outside of just writing it down, and I will need help getting the data on the prospects, but it is available at mlb.com. Further, the data is pretty easy to parse through.

#What time frame do you want/need data from?
Because I am looking at recent rebuilds, I might need data from the last 15 years or so. But with the data we have in baseball archives, this shouldn't be too big of a problem.

#Will you have to improve or augment the data in order to answer your question? If so, how?
I don't believe so. The data I will be working with is fairly straightforward for what I'm interested in.

#Are there other examples of this kind of work (specific or general)? If so, provide them.
https://community.fangraphs.com/an-examination-of-rebuilding-team-timelines/

https://pirates.pittsburghbaseball.com/2021/10/08/how-many-years-should-a-rebuild-take-for-an-mlb-team/?utm_source=rss&utm_medium=rss&utm_campaign=how-many-years-should-a-rebuild-take-for-an-mlb-team

https://www.espn.com/mlb/insider/insider/story/_/id/34407508/when-does-rebuild-fail-detroit-tigers-disaster-means-your-team

#If your data is available and you have it, create code blocks to load the appropriate libraries and then read in the data.
I don't have any of the data yet, but are there any libraries other than tidyverse that I will need to load for this investigation?

##EXPLORATION (10/16/22) - This is the first part of my investigation — looking at the level of talent teams lose during the trade deadline.
What I was interested in doing for this portion of the assignment was trying to see if I could create dataframes that would show which players were traded. I created two key dataframes for each of the years 2021 and 2022. I made one dataframe that showed all the departures throughout the season and one with departures at the trade deadline. 

I wanted to do this because there are some players who might have been waived, or they might have been part of a less important trade. For this investigation, I want to look at major moves teams make when they blow up their rosters, so those marginal trades aren't super interesting. However, I'm still curious as about how different a team looks at the beginning and end of a season.

My plan is to create similar dataframes for other teams that have undergone similar rebuild situations near the trade deadline This will allow me to isolate the talent a team loses when it decides to fire sale its roster at the trade deadline. In other words, it will allow me to summarize the gap in production of who is out and in. 

Doing all this is a way to try to quantify the hole teams put themselves in. I will look to answer questions like How long does it take a team to return to relevancy, and Can we see how long that process takes?

```{r}
#I'm loading the necessary libraries for this set of queries
library(baseballr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{r}
#I'm looking at all the MLB teams
mlb_teams(season = 2022, sport_ids = c(1))

#This is me creating a dataframe for the Nats roster on April 6, 2021 — Opening Day
nats_roster_apr6 <- mlb_rosters(team_id = 120, season = 2021, date = "2021-04-06", roster_type = '40Man')

#This is me creating a dataframe for the Nats roster on October 3, 2021 — Last game of the season
nats_roster_oct3 <- mlb_rosters(team_id = 120, season = 2021, date = "2021-10-03", roster_type = '40Man')

#This is creating a new dataframe for the guys who started the season with the team but left before the end of the year
opening_day_to_last_day_21 <- anti_join(nats_roster_apr6, nats_roster_oct3, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

#This is creating a new dataframe for the guys who didn't start the season with the team but ended the year as a Nat
opening_day_to_last_day_newbies_21 <- anti_join(nats_roster_oct3, nats_roster_apr6, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

#According to opening_day_to_last_day_21, 14 of the players on the 40-man roster left the team, which is 35% of the roster, a solid amount during the season. That was without expiring contracts or players getting cut, which happens in the offseason. This could be interesting information to use in a potential graph, for we could track the change in the team over the course of the season.

#This is me creating a dataframe for the Nats roster on July 20, 2021 — pre-trade deadline
nats_roster_jul20 <- mlb_rosters(team_id = 120, season = 2021, date = "2021-07-20", roster_type = '40Man')

#This is me creating a dataframe for the Nats roster on August 05, 2021 — post-trade deadline
nats_roster_aug5 <- mlb_rosters(team_id = 120, season = 2021, date = "2021-08-05", roster_type = '40Man')

#This is creating a new dataframe for the guys who were traded at the deadline
departures_trade_deadline_21 <- anti_join(nats_roster_jul20, nats_roster_aug5, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

#This is creating a new dataframe for the guys who were acquired at the deadline
acquisitions_trade_deadline_21 <- anti_join(nats_roster_aug5, nats_roster_jul20, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
```

#I just wanted to separate the dataframes for 2021 and 2022 so that it wouldn't be one big jumbled codeblock.
```{r}
#This is me creating a dataframe for the Nats roster on April 7, 2022 — Opening Day
nats_roster_apr7 <- mlb_rosters(team_id = 120, season = 2022, date = "2022-04-07", roster_type = '40Man')

#This is me creating a dataframe for the Nats roster on October 5, 2022 — Last game of the season
nats_roster_oct5 <- mlb_rosters(team_id = 120, season = 2022, date = "2022-10-05", roster_type = '40Man')

#This is creating a new dataframe for the guys who started the season with the team but left before the end of the year
opening_day_to_last_day_22 <- anti_join(nats_roster_apr7, nats_roster_oct5, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

#This is creating a new dataframe for the guys who didn't start the season with the team but ended the year as a Nat
opening_day_to_last_day_newbies_22 <- anti_join(nats_roster_oct5, nats_roster_apr7, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

#According to opening_day_to_last_day_22, 10 of the players on the 40-man roster left the team, which is 25% of the roster. While this is less than the previous year, it is of comparable (if not greater) value because Juan Soto was included in the deal. This could be interesting information to use in a potential graph, for we could track the change in the team over the course of the season.

#This is me creating a dataframe for the Nats roster on July 28, 2022 — pre-trade deadline
nats_roster_aug1 <- mlb_rosters(team_id = 120, season = 2022, date = "2022-07-28", roster_type = '40Man')

#This is me creating a dataframe for the Nats roster on August 10, 2022 — post-trade deadline
nats_roster_aug10 <- mlb_rosters(team_id = 120, season = 2022, date = "2022-08-10", roster_type = '40Man')

#This is creating a new dataframe for the guys who were traded at the deadline
departures_trade_deadline_22 <- anti_join(nats_roster_aug1, nats_roster_aug10, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
  
library(baseballr)
voigt_batter_stats <- mlb_player_game_stats(person_id = 572228)

hassell_minor_league_stats <- fg_milb_batter_game_logs(690987, 2022)

#This is creating a new dataframe for the guys who were acquired at the deadline
acquisitions_trade_deadline_22 <- anti_join(nats_roster_aug10, nats_roster_aug1, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
```

#I wanted to see how things would change if I changed the roster_typ from '40man' to 'fullRoster.' Just using '40man' doesn't account for some of the prospects that were acquired in trades; it only gives us players who debuted for the team or were on the 40-man roster at some point. But, this investigation is looking at prospects acquired in trades, so I just have to figure out the best way to create a dataframe with impact players — some of the guys on the lists that we create below will never play at the pro level. 

#I will expand this portion of the project for the next deliverable, but I think it is also important to use '40man' as part of my investigation. I just want to think a bit more on the best way to get the departed and acquired players on teams, but I wanted to show you that I have kicked around the idea of using fullRoster as well.
```{r}

nats_roster_jul26 <- mlb_rosters(team_id = 120, season = 2022, date = "2022-07-26", roster_type = 'fullRoster')


nats_roster_aug9 <- mlb_rosters(team_id = 120, season = 2022, date = "2022-08-09", roster_type = 'fullRoster')


departures_trade_deadline_22_test <- anti_join(nats_roster_jul26, nats_roster_aug9, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))


acquisitions_trade_deadline_22_test <- anti_join(nats_roster_aug9, nats_roster_jul26, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
```

#Refining (10/16/22) — In this portion of the assignment, I'm going to try and chart the record of the Orioles and Marlins, two teams that traded away big-name players at the trade deadline.
```{r}
#I'm looking at all the MLB teams
mlb_teams(season = 2022, sport_ids = c(1))
```

#Orioles information
```{r}
#Dataframe for the O's 2019 record 
os_2019_record <- bref_team_results('BAL', 2019) %>%
  filter(Gm == '162')

#Dataframe for the O's 2020 record 
os_2020_record <- bref_team_results('BAL', 2020) %>%
  filter(Gm == '60')

#Dataframe for the O's 2021 record 
os_2021_record <- bref_team_results('BAL', 2021) %>%
  filter(Gm == '162')

#Dataframe for the O's 2022 record 
os_2022_record <- bref_team_results('BAL', 2022) %>%
  filter(Gm == '162')

#Dataframe for O's four-year record
os_record_19_22 <- bind_rows(os_2019_record, os_2020_record, os_2021_record, os_2022_record)
  
#Separating the Record column into Wins and Losses
os_record_wins_losses_separate <- separate(os_record_19_22, col=Record, into=c('Wins', 'Losses'), sep='-') %>%
  mutate(Wins = as.numeric(Wins), Losses = as.numeric(Losses)) %>%
  mutate(win_percentage = (Wins/(Wins+Losses)*100))

#Graphing the wins by year using a line chart
ggplot(os_record_wins_losses_separate, aes(x=Year, y=win_percentage, group=Tm)) +         
  geom_line() +
  geom_point()
```
#At the 2018 trade deadline, the Orioles traded away Darren O’Day, Kevin Gausman, Manny machado, Zach Britton and Jonathan Schoop. They won 54 the next season before winning 52 the following year. But in 2022, the team jumped up to 83 wins. Using just wins, it looks like the selloff the O's undertook in 2018 is paying off. Again, one of my criteria for looking at whether a team's selloff works is how many years it takes before the team has a winning record, so it took the O's three and a half seasons to log do so.

#I realize now that the 25 wins in 2020 skews this graph. It makes it look like the team in improving, when in reality that's 25 wins out of 60 games. So over a full season, that would equate to about 68 wins, so our graph would look different. I will try and correct this so the graph isnt as misleading.

#Marlins Information
```{r}
#Dataframe for the Fins' 2018 record 
fins_2018_record <- bref_team_results('MIA', 2018) %>%
    filter(Gm == '161')

#Dataframe for the Fins' 2019 record 
fins_2019_record <- bref_team_results('MIA', 2019) %>%
    filter(Gm == '162')

#Dataframe for the Fins' 2020 record 
fins_2020_record <- bref_team_results('MIA', 2020) %>%
    filter(Gm == '60')

#Dataframe for the Fins' 2021 record 
fins_2021_record <- bref_team_results('MIA', 2021) %>%
    filter(Gm == '162')

#Dataframe for the Fins' 2022 record 
fins_2022_record <- bref_team_results('MIA', 2022) %>%
  filter(Gm == '162')

#Dataframe for Fins' five-year record
fins_record_18_22 <- bind_rows(fins_2018_record, fins_2019_record, fins_2020_record, fins_2021_record, fins_2022_record)

#Separating the Record column into Wins and Losses
fins_record_wins_losses_separate <- separate(fins_record_18_22, col=Record, into=c('Wins', 'Losses'), sep='-') %>%
  mutate(Wins = as.numeric(Wins), Losses = as.numeric(Losses)) %>%
  mutate(win_percentage = (Wins/(Wins+Losses)*100))

#Graphing the wins by year using a line chart
ggplot(fins_record_wins_losses_separate, aes(x=Year, y=win_percentage, group=Tm)) +         
  geom_line() +
  geom_point()
```
#In the 2018 offseason, the Marlins traded away Dee Gordon, Christian Yelich, Giancarlo Stanton and Marcel Ozuna. They won 63 the next season before winning 57 the following year. But in 2021 and then again in 2022, the team jumped up from 67 to 69 wins. Now, this is still a pretty eh baseball team, but it is winning more games. Using just wins, it looks like the selloff the Marlins undertook in 2018 is not paying off yet. Again, one of my criteria for looking at whether a team's selloff works is how many years it takes before the team has a winning record, so we're still waiting on the Marlins here.

#I realize now that the 31 wins in 2020 skews this graph. It makes it look like the team in improving, when in reality that's 31 wins out of 60 games. So over a full season, that would equate to about 84 wins, so our graph would look different. I will try and correct this so the graph isnt as misleading.



#Notes to myself
Stats

cope data from baseball reference, share&export —> get table as cvs

in r studio, files, new blank file, text file, give it a title with .csv paste in contents and hit save
copy and put into 

bind_rows function to get 2022 stats for player

is it worth exploring not just the percentage of the roster but percentage of various kinds of production that is lost as well? Not as a replacement for roster comparisons but an addition to them. I agree that 40-man roster probably is better to use for calculating actual gains and losses of players, and am interested to see what you find about minor league players, since you'll need to include their minor league production in your analysis.

#More Refining — I am thinking about looking at different information for my project. I realize that the Covid year kind of skews what I am looking at, even if I adjust for win percentage and not overall wins, so instead I'm going to look at how long it takes teams to get back to a world series
```{r}
#I'm looking at all the MLB teams
mlb_teams(season = 2022, sport_ids = c(1))
```

```{r}
braves_1995_record <- bref_team_results('ATL', 1995) %>%
    filter(Gm == '144')
braves_1996_record <- bref_team_results('ATL', 1996) %>%
    filter(Gm == '162')
braves_1997_record <- bref_team_results('ATL', 1997) %>%
    filter(Gm == '162')
braves_1998_record <- bref_team_results('ATL', 1998) %>%
    filter(Gm == '162')
braves_1999_record <- bref_team_results('ATL', 1999) %>%
    filter(Gm == '162')
braves_2000_record <- bref_team_results('ATL', 2000) %>%
    filter(Gm == '162')
braves_2001_record <- bref_team_results('ATL', 2001) %>%
    filter(Gm == '162')
braves_2002_record <- bref_team_results('ATL', 2002) %>%
    filter(Gm == '161')
braves_2003_record <- bref_team_results('ATL', 2003) %>%
    filter(Gm == '162')
braves_2004_record <- bref_team_results('ATL', 2004) %>%
    filter(Gm == '162')
braves_2005_record <- bref_team_results('ATL', 2005) %>%
    filter(Gm == '162')
braves_2006_record <- bref_team_results('ATL', 2006) %>%
    filter(Gm == '162')
braves_2007_record <- bref_team_results('ATL', 2007) %>%
    filter(Gm == '162')
braves_2008_record <- bref_team_results('ATL', 2008) %>%
    filter(Gm == '162')
braves_2009_record <- bref_team_results('ATL', 2009) %>%
    filter(Gm == '162')
braves_2010_record <- bref_team_results('ATL', 2010) %>%
    filter(Gm == '162')
braves_2011_record <- bref_team_results('ATL', 2011) %>%
    filter(Gm == '162')
braves_2012_record <- bref_team_results('ATL', 2012) %>%
    filter(Gm == '162')
braves_2013_record <- bref_team_results('ATL', 2013) %>%
    filter(Gm == '162')
braves_2014_record <- bref_team_results('ATL', 2014) %>%
    filter(Gm == '162')
braves_2015_record <- bref_team_results('ATL', 2015) %>%
    filter(Gm == '162')
braves_2016_record <- bref_team_results('ATL', 2016) %>%
    filter(Gm == '161')
braves_2017_record <- bref_team_results('ATL', 2017) %>%
    filter(Gm == '162')
braves_2018_record <- bref_team_results('ATL', 2018) %>%
    filter(Gm == '162')
braves_2019_record <- bref_team_results('ATL', 2019) %>%
    filter(Gm == '162')
braves_2020_record <- bref_team_results('ATL', 2020) %>%
    filter(Gm == '60')
braves_2021_record <- bref_team_results('ATL', 2021) %>%
    filter(Gm == '161')

braves_record_95_21 <- bind_rows(braves_1995_record, braves_1996_record, braves_1997_record, braves_1998_record, braves_1999_record, braves_2000_record, braves_2001_record, braves_2002_record, braves_2003_record, braves_2004_record, braves_2005_record, braves_2006_record, braves_2007_record, braves_2008_record, braves_2009_record, braves_2010_record, braves_2011_record, braves_2012_record, braves_2013_record, braves_2014_record, braves_2015_record, braves_2016_record, braves_2017_record, braves_2018_record, braves_2019_record, braves_2020_record, braves_2021_record)

#Separating the Record column into Wins and Losses
braves_record_wins_losses_separate <- separate(braves_record_95_21, col=Record, into=c('Wins', 'Losses'), sep='-') %>%
  mutate(Wins = as.numeric(Wins), Losses = as.numeric(Losses)) %>%
  mutate(win_percentage = (Wins/(Wins+Losses)*100))

#Graphing the wins by year using a line chart
ggplot(braves_record_wins_losses_separate, aes(x=Year, y=win_percentage, group=Tm)) +         
  geom_line() +
  geom_point()
```

#Player change since the Nats won the world series (2019)
```{r}
nats_od_roster_2019 <- mlb_rosters(team_id = 120, season = 2019, date = "2019-03-28", roster_type = 'fullSeason')

nats_od_roster_2020 <- mlb_rosters(team_id = 120, season = 2020, date = "2020-07-23", roster_type = 'fullSeason')

nats_od_roster_2021 <- mlb_rosters(team_id = 120, season = 2021, date = "2021-04-06", roster_type = 'fullSeason')

nats_od_roster_2022 <- mlb_rosters(team_id = 120, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')

nats_od_19_to_20 <- anti_join(nats_od_roster_2019, nats_od_roster_2020, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

nats_od_20_to_21 <- anti_join(nats_od_roster_2020, nats_od_roster_2021, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

nats_od_21_to_22 <- anti_join(nats_od_roster_2021, nats_od_roster_2022, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

nats_19_20_departures <- nats_od_19_to_20 %>% group_by(season) %>% summarize(departures = n()) %>% mutate(full_season = nats_od_roster_2019 %>% nrow()) %>% mutate(pct_departed = departures/full_season)

nats_20_21_departures <- nats_od_20_to_21 %>% group_by(season) %>% summarize(departures = n()) %>% mutate(full_season = nats_od_roster_2019 %>% nrow()) %>% mutate(pct_departed = departures/full_season)

nats_21_22_departures <- nats_od_21_to_22 %>% group_by(season) %>% summarize(departures = n()) %>% mutate(full_season = nats_od_roster_2019 %>% nrow()) %>% mutate(pct_departed = departures/full_season)

nats_total_players_lost <- bind_rows(nats_19_20_departures, nats_20_21_departures, nats_21_22_departures) %>%
  rename("Year" = "season")

nats_players_lost_plus_win_percentage <- inner_join(nats_total_players_lost, nats_record_wins_losses_separate, by = "Year")

ggplot() + geom_point(data=nats_players_lost_plus_win_percentage, aes(x=win_percentage, y=pct_departed))

#bind_rows

#Nats wins
nats_2019_record <- bref_team_results('WSN', 2019) %>%
    filter(Gm == '162')

nats_2020_record <- bref_team_results('WSN', 2020) %>%
    filter(Gm == '60')

nats_2021_record <- bref_team_results('WSN', 2021) %>%
    filter(Gm == '162')

nats_2022_record <- bref_team_results('WSN', 2022) %>%
    filter(Gm == '162')

nats_record_19_22 <- bind_rows(nats_2019_record, nats_2020_record, nats_2021_record, nats_2022_record)

#Separating the Record column into Wins and Losses
nats_record_wins_losses_separate <- separate(nats_record_19_22, col=Record, into=c('Wins', 'Losses'), sep='-') %>%
  mutate(Wins = as.numeric(Wins), Losses = as.numeric(Losses)) %>%
  mutate(win_percentage = (Wins/(Wins+Losses)*100)) %>% select(Year, win_percentage)

nats_players_lost <- bind_rows(nats_od_19_to_20, nats_od_20_to_21, nats_od_21_to_22)

#Graphing the wins by year using a line chart
ggplot(nats_record_wins_losses_separate, aes(x=Year, y=win_percentage, group=Tm)) +         
  geom_line() +
  geom_point()
```

#Player change since the Red Sox won the world series (2018)
```{r}
rsox_od_roster_2018 <- mlb_rosters(team_id = 111, season = 2018, date = "2018-04-05", roster_type = 'fullSeason')

rsox_od_roster_2019 <- mlb_rosters(team_id = 111, season = 2019, date = "2019-04-09", roster_type = 'fullSeason')

rsox_od_roster_2020 <- mlb_rosters(team_id = 111, season = 2020, date = "2020-07-24", roster_type = 'fullSeason')

rsox_od_roster_2021 <- mlb_rosters(team_id = 111, season = 2021, date = "2021-04-02", roster_type = 'fullSeason')

rsox_od_roster_2022 <- mlb_rosters(team_id = 111, season = 2022, date = "2022-04-15", roster_type = 'fullSeason')

rsox_od_18_to_19 <- anti_join(rsox_od_roster_2018, rsox_od_roster_2019, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

rsox_od_19_to_20 <- anti_join(rsox_od_roster_2019, rsox_od_roster_2020, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

rsox_od_20_to_21 <- anti_join(rsox_od_roster_2020, rsox_od_roster_2021, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

rsox_od_21_to_22 <- anti_join(rsox_od_roster_2021, rsox_od_roster_2022, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
```

#Player change since the Astros won the world series (2017)
```{r}
astros_od_roster_2017 <- mlb_rosters(team_id = 117, season = 2017, date = "2017-04-03", roster_type = 'fullSeason')

astros_od_roster_2018 <- mlb_rosters(team_id = 117, season = 2018, date = "2018-04-02", roster_type = 'fullSeason')

astros_od_roster_2019 <- mlb_rosters(team_id = 117, season = 2019, date = "2019-04-05", roster_type = 'fullSeason')

astros_od_roster_2020 <- mlb_rosters(team_id = 117, season = 2020, date = "2020-07-24", roster_type = 'fullSeason')

astros_od_roster_2021 <- mlb_rosters(team_id = 117, season = 2021, date = "2021-04-08", roster_type = 'fullSeason')

astros_od_roster_2022 <- mlb_rosters(team_id = 117, season = 2022, date = "2022-04-18", roster_type = 'fullSeason')

astros_od_17_to_18 <- anti_join(astros_od_roster_2017, astros_od_roster_2018, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

astros_od_18_to_19 <- anti_join(astros_od_roster_2018, astros_od_roster_2019, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

astros_od_19_to_20 <- anti_join(astros_od_roster_2019, astros_od_roster_2020, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

astros_od_20_to_21 <- anti_join(astros_od_roster_2020, astros_od_roster_2021, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

astros_od_21_to_22 <- anti_join(astros_od_roster_2021, astros_od_roster_2022, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
```

#Player change since the Cubs won the world series (2016)
```{r}
cubs_od_roster_2016 <- mlb_rosters(team_id = 112, season = 2016, date = "2016-04-11", roster_type = 'fullSeason')

cubs_od_roster_2017 <- mlb_rosters(team_id = 112, season = 2017, date = "2017-04-10", roster_type = 'fullSeason')

cubs_od_roster_2018 <- mlb_rosters(team_id = 112, season = 2018, date = "2018-04-10", roster_type = 'fullSeason')

cubs_od_roster_2019 <- mlb_rosters(team_id = 112, season = 2019, date = "2019-04-08", roster_type = 'fullSeason')

cubs_od_roster_2020 <- mlb_rosters(team_id = 112, season = 2020, date = "2020-07-24", roster_type = 'fullSeason')

cubs_od_roster_2021 <- mlb_rosters(team_id = 112, season = 2021, date = "2021-04-01", roster_type = 'fullSeason')

cubs_od_roster_2022 <- mlb_rosters(team_id = 112, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')

cubs_od_16_to_17 <- anti_join(cubs_od_roster_2016, cubs_od_roster_2017, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

cubs_od_17_to_18 <- anti_join(cubs_od_roster_2017, cubs_od_roster_2018, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

cubs_od_18_to_19 <- anti_join(cubs_od_roster_2018, cubs_od_roster_2019, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

cubs_od_19_to_20 <- anti_join(cubs_od_roster_2019, cubs_od_roster_2020, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

cubs_od_20_to_21 <- anti_join(cubs_od_roster_2020, cubs_od_roster_2021, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

cubs_od_21_to_22 <- anti_join(cubs_od_roster_2021, cubs_od_roster_2022, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
```

#Player change since the Royals won the world series (2015)
```{r}
royals_od_roster_2015 <- mlb_rosters(team_id = 118, season = 2015, date = "2015-04-06", roster_type = 'fullSeason')

royals_od_roster_2016 <- mlb_rosters(team_id = 118, season = 2016, date = "2016-04-03", roster_type = 'fullSeason')

royals_od_roster_2017 <- mlb_rosters(team_id = 118, season = 2017, date = "2017-04-10", roster_type = 'fullSeason')

royals_od_roster_2018 <- mlb_rosters(team_id = 118, season = 2018, date = "2018-03-29", roster_type = 'fullSeason')

royals_od_roster_2019 <- mlb_rosters(team_id = 118, season = 2019, date = "2019-03-28", roster_type = 'fullSeason')

royals_od_roster_2020 <- mlb_rosters(team_id = 118, season = 2020, date = "2020-07-31", roster_type = 'fullSeason')

royals_od_roster_2021 <- mlb_rosters(team_id = 118, season = 2021, date = "2021-04-01", roster_type = 'fullSeason')

royals_od_roster_2022 <- mlb_rosters(team_id = 118, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')

royals_od_15_to_16 <- anti_join(royals_od_roster_2015, royals_od_roster_2016, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

royals_od_16_to_17 <- anti_join(royals_od_roster_2015, royals_od_roster_2016, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

royals_od_17_to_18 <- anti_join(royals_od_roster_2015, royals_od_roster_2016, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

royals_od_18_to_19 <- anti_join(royals_od_roster_2015, royals_od_roster_2016, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

royals_od_19_to_20 <- anti_join(royals_od_roster_2015, royals_od_roster_2016, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

royals_od_20_to_21 <- anti_join(royals_od_roster_2015, royals_od_roster_2016, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

royals_od_21_to_22 <- anti_join(royals_od_roster_2015, royals_od_roster_2016, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
```

