---
title: "presentation.Rmd"
output: html_document
date: "2022-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#I'm loading the necessary libraries for this set of queries
library(baseballr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{r}
#I'm looking at all the MLB teams
mlb_teams(season = 2022, sport_ids = c(1))
```

#Teams that "blew up" their rosters
```{r}
#Every player who logged a major-league game for the Nats in 2022
nats_od_roster_2022 <- mlb_rosters(team_id = 120, season = 2022, date = "2021-04-07", roster_type = 'fullSeason')

#Every player who logged a major-league game for the A's in 2022
athletics_od_roster_2022 <- mlb_rosters(team_id = 133, season = 2022, date = "2022-04-08", roster_type = 'fullSeason')

#Every player who logged a major-league game for the Reds in 2022
reds_od_roster_2022 <- mlb_rosters(team_id = 113, season = 2022, date = "2022-04-08", roster_type = 'fullSeason')
```

#"Sellers" at the trade deadline
```{r}
nats_departures <- nats_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Nats')

athletics_departures <- athletics_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'As')

reds_departures <- reds_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Reds')
```

```{r}
#Combined sellers dataframe
combined_sellers <- bind_rows(nats_departures, athletics_departures, reds_departures)
```

```{r}
#Stacked bar chart showing the player status descriptions for the Nats, A's and Reds
ggplot() + 
  geom_bar(data=combined_sellers, aes(x=team, weight=count, fill=status_description)) +
  coord_flip() +
   labs(title="What does 'blowing up' a roster actually mean in Major League Baseball?", subtitle="The Nationals traded fewer players than other 'rebuilding' teams but had the league's worst record.", caption="Source: Baseball-Reference | By Ian Decker")
```

```{r}
nats_od_roster_2021 <- mlb_rosters(team_id = 120, season = 2021, date = "2021-04-06", roster_type = 'fullSeason')

nats_departures_2021 <- nats_od_roster_2021 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Nats')

ggplot() + 
  geom_bar(data=nats_departures_2021, aes(x=team, weight=count, fill=status_description)) +
  coord_flip() +
   labs(title="What does 'blowing up' a roster actually mean in Major League Baseball?", subtitle="The Nationals traded fewer players than other 'rebuilding' teams but had the league's worst record.", caption="Source: Baseball-Reference | By Ian Decker")
```


#AL teams expected to compete for the World Series
```{r}
#Every player who logged a major-league game for the Astros in 2022
astros_od_roster_2022 <- mlb_rosters(team_id = 117, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')

#Every player who logged a major-league game for the Yankees in 2022
yankees_od_roster_2022 <- mlb_rosters(team_id = 147, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')
```

#NL Teams expected to compete for the World Series 
```{r}
#Every player who logged a major-league game for the Dodgers in 2022
dodgers_od_roster_2022 <- mlb_rosters(team_id = 119, season = 2022, date = "2022-04-08", roster_type = 'fullSeason')

#Every player who logged a major-league game for the Mets in 2022
mets_od_roster_2022 <- mlb_rosters(team_id = 121, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')
```

#"Buyers" at the trade deadline
```{r}
astros_departures <- astros_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Astros')

yankees_departures <- yankees_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Yankees')

dodgers_departures <- dodgers_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Dodgers')

mets_departures <- mets_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Mets')
```

```{r}
#Combined buyers dataframe
combined_buyers <- bind_rows(astros_departures, yankees_departures, dodgers_departures, mets_departures)
```

```{r}
#Stacked bar chart showing the player status descriptions for the Astros, Yankees, Dodgers and Mets
ggplot() + 
  geom_bar(data=combined_buyers, aes(x=team, weight=count, fill=status_description)) +
  coord_flip() +
   labs(title="Do teams with World Series hopes 'blow up' their rosters?", subtitle="Most MLB trade deadline buyers alter their rosters midway through the season.", caption="Source: Baseball-Reference | By Ian Decker")
```

```{r}
nats_2005_record <- bref_team_results('WSN', 2005) %>%
    filter(Gm == '162')
nats_2006_record <- bref_team_results('WSN', 2006) %>%
    filter(Gm == '162')
nats_2007_record <- bref_team_results('WSN', 2007) %>%
    filter(Gm == '162')
nats_2008_record <- bref_team_results('WSN', 2008) %>%
    filter(Gm == '161')
nats_2009_record <- bref_team_results('WSN', 2009) %>%
    filter(Gm == '162')
nats_2010_record <- bref_team_results('WSN', 2010) %>%
    filter(Gm == '162')
nats_2011_record <- bref_team_results('WSN', 2011) %>%
    filter(Gm == '161')
nats_2012_record <- bref_team_results('WSN', 2012) %>%
    filter(Gm == '162')
nats_2013_record <- bref_team_results('WSN', 2013) %>%
    filter(Gm == '162')
nats_2014_record <- bref_team_results('WSN', 2014) %>%
    filter(Gm == '162')
nats_2015_record <- bref_team_results('WSN', 2015) %>%
    filter(Gm == '162')
nats_2016_record <- bref_team_results('WSN', 2016) %>%
    filter(Gm == '161')
nats_2017_record <- bref_team_results('WSN', 2017) %>%
    filter(Gm == '162')
nats_2018_record <- bref_team_results('WSN', 2018) %>%
    filter(Gm == '162')
nats_2019_record <- bref_team_results('WSN', 2019) %>%
    filter(Gm == '162')
nats_2020_record <- bref_team_results('WSN', 2020) %>%
    filter(Gm == '60')
nats_2021_record <- bref_team_results('WSN', 2021) %>%
    filter(Gm == '162')
nats_2022_record <- bref_team_results('WSN', 2022) %>%
    filter(Gm == '162')  

nats_overall_record <- bind_rows(nats_2005_record, nats_2006_record, nats_2007_record, nats_2008_record,nats_2009_record, nats_2010_record, nats_2011_record, nats_2012_record, nats_2013_record, nats_2014_record, nats_2015_record, nats_2016_record, nats_2017_record, nats_2018_record, nats_2019_record, nats_2020_record, nats_2021_record, nats_2022_record)

#Separating the Record column into Wins and Losses
nats_record_wins_losses_separate <- separate(nats_overall_record, col=Record, into=c('Wins', 'Losses'), sep='-') %>%
  mutate(Wins = as.numeric(Wins), Losses = as.numeric(Losses)) %>%
  mutate(win_percentage = (Wins/(Wins+Losses)*100))

#Graphing the wins by year using a line chart
ggplot(nats_record_wins_losses_separate, aes(x=Year, y=win_percentage, group=Tm)) +         
  geom_line() +
  geom_point(colour='red') +
   labs(title="Have the Washington Nationals ever been this bad?", subtitle="2022 was the worst season in Washington baseball history, but how long will the rebuild take?", caption="Source: Baseball-Reference | By Ian Decker")
```

#Narrative
When the Washington Nationals traded away Juan Soto in on August 2, they signalled to the fans and the rest of the league that they were 'blowing up' the roster and planning for the future.

But the trade of Soto, and first baseman Josh Bell, was the latest example of an MLB franchise trading away a superstar whom they couldn't afford to keep for younger, and more importantly, cheaper players. 

Looking at the fall from grace Washington baseball has experienced since winnig the World Series in 2019, I was curious as to what 'blowing up' a roster actually meant. Yes, the Nationals traded away a generational talent in Soto and a former all-star in Bell, but aside from a little-remebered trade where they swapped fringe players, the team traded only Soto and Bell. 

A season prior, the team parted ways with eight veteran players, almost thrice as many compared with 2022. From a raw numbers perpesctive, the team detonated its roster to a larger degree in 2021 than it did a year later in 2022. This led to the question that drove my analysis: Do we have an inflated sense of roster blowing up?


What happens to players when they depart


as for the second question, i think you should briefly summarize what your original idea was, the obstacles you encountered and then how you got to where you are now. If there's some code that helps illustrate any of those points, feel free to use it as an example.

I thought at first I'd look at X, but quickly realized that I couldn't due to not having the data I needed.

and contain explanatory sections that describe your code and the choices you made. Explain the methodological choices you made and why you made them.  Organize the narrative like a story - what did you find (or not find) and why readers should care.


