---
title: "presentation.Rmd"
output: html_document
date: "2022-10-31"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#I'm loading the necessary libraries for this set of queries
library(baseballr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

```{r}
#I'm looking at all the MLB teams
mlb_teams(season = 2022, sport_ids = c(1))
```

#Teams that "blew up" their rosters
```{r}
#Every player who logged a major-league game for the Nats in 2022
nats_od_roster_2022 <- mlb_rosters(team_id = 120, season = 2022, date = "2021-04-07", roster_type = 'fullSeason')

#Every player who logged a major-league game for the A's in 2022
athletics_od_roster_2022 <- mlb_rosters(team_id = 133, season = 2022, date = "2022-04-08", roster_type = 'fullSeason')

#Every player who logged a major-league game for the Reds in 2022
reds_od_roster_2022 <- mlb_rosters(team_id = 113, season = 2022, date = "2022-04-08", roster_type = 'fullSeason')
```

#"Sellers" at the trade deadline
```{r}
nats_departures <- nats_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Nats')

athletics_departures <- athletics_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'As')

reds_departures <- reds_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Reds')
```

```{r}
#Combined sellers dataframe
combined_sellers <- bind_rows(nats_departures, athletics_departures, reds_departures)
```

```{r}
#Stacked bar chart showing the player status descriptions for the Nats, A's and Reds
ggplot() + 
  geom_bar(data=combined_sellers, aes(x=team, weight=count, fill=status_description)) +
  coord_flip() +
   labs(title="Graph 1A: What does 'blowing up' a roster actually mean in Major League Baseball?", subtitle="The Nationals traded fewer players than other 'rebuilding' teams but had the league's worst record.", caption="Source: Baseball-Reference | By Ian Decker")
```

#Looking at how the Nats 'blew up' their roster in 2021
```{r}
nats_od_roster_2021 <- mlb_rosters(team_id = 120, season = 2021, date = "2021-04-06", roster_type = 'fullSeason')

nats_departures_2021 <- nats_od_roster_2021 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Nats')

ggplot() + 
  geom_bar(data=nats_departures_2021, aes(x=team, weight=count, fill=status_description)) +
  coord_flip() +
   labs(title="Graph 1B: What does 'blowing up' a roster actually mean in Major League Baseball?", subtitle="The Nationals traded fewer players than other 'rebuilding' teams but had the league's worst record.", caption="Source: Baseball-Reference | By Ian Decker")
```

#AL teams expected to compete for the World Series
```{r}
#Every player who logged a major-league game for the Astros in 2022
astros_od_roster_2022 <- mlb_rosters(team_id = 117, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')

#Every player who logged a major-league game for the Yankees in 2022
yankees_od_roster_2022 <- mlb_rosters(team_id = 147, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')
```

#NL Teams expected to compete for the World Series 
```{r}
#Every player who logged a major-league game for the Dodgers in 2022
dodgers_od_roster_2022 <- mlb_rosters(team_id = 119, season = 2022, date = "2022-04-08", roster_type = 'fullSeason')

#Every player who logged a major-league game for the Mets in 2022
mets_od_roster_2022 <- mlb_rosters(team_id = 121, season = 2022, date = "2022-04-07", roster_type = 'fullSeason')
```

#"Buyers" at the trade deadline
```{r}
astros_departures <- astros_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Astros')

yankees_departures <- yankees_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Yankees')

dodgers_departures <- dodgers_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Dodgers')

mets_departures <- mets_od_roster_2022 %>%
  group_by(status_description) %>%
  summarise(count = n()) %>% 
  mutate(team = 'Mets')
```

```{r}
#Combined buyers dataframe
combined_buyers <- bind_rows(astros_departures, yankees_departures, dodgers_departures, mets_departures)
```

```{r}
#Stacked bar chart showing the player status descriptions for the Astros, Yankees, Dodgers and Mets
ggplot() + 
  geom_bar(data=combined_buyers, aes(x=team, weight=count, fill=status_description)) +
  coord_flip() +
   labs(title="Graph 2: Do teams with World Series hopes 'blow up' their rosters?", subtitle="Most MLB trade deadline buyers alter their rosters midway through the season.", caption="Source: Baseball-Reference | By Ian Decker")
```

```{r}
#Nats year-by-year records 2005-2022
nats_2005_record <- bref_team_results('WSN', 2005) %>%
    filter(Gm == '162')
nats_2006_record <- bref_team_results('WSN', 2006) %>%
    filter(Gm == '162')
nats_2007_record <- bref_team_results('WSN', 2007) %>%
    filter(Gm == '162')
nats_2008_record <- bref_team_results('WSN', 2008) %>%
    filter(Gm == '161')
nats_2009_record <- bref_team_results('WSN', 2009) %>%
    filter(Gm == '162')
nats_2010_record <- bref_team_results('WSN', 2010) %>%
    filter(Gm == '162')
nats_2011_record <- bref_team_results('WSN', 2011) %>%
    filter(Gm == '161')
nats_2012_record <- bref_team_results('WSN', 2012) %>%
    filter(Gm == '162')
nats_2013_record <- bref_team_results('WSN', 2013) %>%
    filter(Gm == '162')
nats_2014_record <- bref_team_results('WSN', 2014) %>%
    filter(Gm == '162')
nats_2015_record <- bref_team_results('WSN', 2015) %>%
    filter(Gm == '162')
nats_2016_record <- bref_team_results('WSN', 2016) %>%
    filter(Gm == '161')
nats_2017_record <- bref_team_results('WSN', 2017) %>%
    filter(Gm == '162')
nats_2018_record <- bref_team_results('WSN', 2018) %>%
    filter(Gm == '162')
nats_2019_record <- bref_team_results('WSN', 2019) %>%
    filter(Gm == '162')
nats_2020_record <- bref_team_results('WSN', 2020) %>%
    filter(Gm == '60')
nats_2021_record <- bref_team_results('WSN', 2021) %>%
    filter(Gm == '162')
nats_2022_record <- bref_team_results('WSN', 2022) %>%
    filter(Gm == '162')  

#Single dataframe with Nats overall record
nats_overall_record <- bind_rows(nats_2005_record, nats_2006_record, nats_2007_record, nats_2008_record,nats_2009_record, nats_2010_record, nats_2011_record, nats_2012_record, nats_2013_record, nats_2014_record, nats_2015_record, nats_2016_record, nats_2017_record, nats_2018_record, nats_2019_record, nats_2020_record, nats_2021_record, nats_2022_record)

#Separating the Record column into Wins and Losses
nats_record_wins_losses_separate <- separate(nats_overall_record, col=Record, into=c('Wins', 'Losses'), sep='-') %>%
  mutate(Wins = as.numeric(Wins), Losses = as.numeric(Losses)) %>%
  mutate(win_percentage = (Wins/(Wins+Losses)*100))

#Graphing the wins by year using a line chart
ggplot(nats_record_wins_losses_separate, aes(x=Year, y=win_percentage, group=Tm)) +         
  geom_line() +
  geom_point(colour='red') +
   labs(title="Graph 3: Have the Washington Nationals ever been this bad?", subtitle="2022 was the worst season in Washington baseball history, but how long will the rebuild take?", caption="Source: Baseball-Reference | By Ian Decker")
```

#Narrative
When the Washington Nationals traded away Juan Soto in on August 2, they signalled to the fans and the rest of the league that they were 'blowing up' the roster and planning for the future.

But the trade of Soto, and first baseman Josh Bell, was the latest example of an MLB franchise trading away a superstar whom they couldn't afford to keep for younger, and more importantly, cheaper players. 

Looking at the fall from grace Washington baseball has experienced since winnig the World Series in 2019, I was curious as to what 'blowing up' a roster actually meant. Yes, the Nationals traded away a generational talent in Soto and a former all-star in Bell, but aside from a little-remebered trade where they swapped fringe players, the team traded only Soto and Bell. 

A season prior, the team parted ways with eight veteran players, almost thrice as many compared with 2022. From a raw numbers perpesctive, the team detonated its roster to a larger degree in 2021 than it did a year later in 2022. This led to the question that drove my analysis: Do we have an inflated sense of roster blowing up?

In 2021, the Nationals netted 12 players from the eight they traded away, and seven for three in 2022, but the caliber of player is what really gives us a sense of what a team does to its roster. It is easy to think that the 2022 rendition of the Nats decimated its roster, but the 2021 team also traded several big-name players, but we think more about the Soto trade. Why? We perceive roster deconstructions differently. 

For instance, looking at opening day rosters from seven major league teams (the Washington Nationals, Oakland Athletics, Cincinnati Reds, Houston Astros, NY Yankees, LA Dodgers and NY Mets) I wanted to look at whether even contending teams 'blow up their rosters.' For my investigation, I chose three teams (Nats, A's and Reds) who acted as sellers at the trade deadline, sending proven big-leaguers for prospects. I also looked at the roster moves of four teams expected to compete for a title this year to see what 'blowing up' a roster actually meant.

The results showed that this is an extremely nuanced question, but it did shed light on debunking our understanding of a team hitting reset on its roster. For example, most analysts agree that the Nationals made the biggest deadline day move when they traded Juan Soto, making it the team that blew up its roster to the largest degree. However, in Graph 1A, the Nats (three trades) made fewer moves — reset their roster to a lesser degree — than both the A's (four) and Reds (seven). 

What's more, Washington (eight) carried fewer minor league contracts — generally associated with a losing team — than Oakland (10) and Cincy (nine). The logic is as follows: the more minor league-level players a major league roster has, the worse that team will perform, thus signifying that team is in rebuild mode.

I also wanted to compare Washington's trade deadlines from the past two seasons. If we compare the Nats' moves in Graph 1A and 1B, we see that the fuchsia bar (trades) is larger in 2021 than in 2022. So according to our chart, the Nats blew up their roster to a larger degree in '21 than '22, but we know that's not true from both a talent and a player return standpoint. And while it is clear that Washington has embraced a rebuild just three years after it one the World Series, do contending teams also 'blow up' their rosters?

The Houston Astros have made the American League Championship Series every year since 2017 and have been one of the most dominant teams in baseball; The New York Yankees have the third-largest payroll in the bigs; the Los Angeles Dodgers (highest payroll) contend on a yearly basis; the New York Mets (second-highest payroll) shelled out massive contracts in the offseason to compete for a title — nobody would think these are teams blowing up their rosters. 

But, when we look at Graph 2, we see that each team altered its roster throughout the season, and the Astros (two trades) were the only team that made fewer trades than the Nats. The Mets also made three trades, the Dodgers had four and the Yankees made five. Now, this evaluation does not look at the statistical or economic value of the players involved in the trades, but it does, however, illustrate that every team 'blows up' its roster — to an extent — during the season; the difference is what a team trades away (fringe players or all-stars) and what it gets in return (win-now players or prospects for the future). 

When I started this investigation, I wanted to look at when teams decide to blow up their rosters. I tried to answer this question by looking at recent sell-offs and what the records of those teams were. The hope was to try and quantify if there's a point when teams decide to blow up their rosters. I was also interested in analyzing how many years it takes before a team is good again — when a losing team either has its first winning record or makes the postseason. 

While this idea didn't fully materialize, I was able to chart Washington's record since the franchise's first season in 2005. This let me look at how long it took the Nats to become a competitive team in the early 2010s. After going 81-81 in 2005, Washington had six straight losing seasons before finally winning more than 50% of its games in 2012, a trend it continued through its World Series-winning season in 2019. Using the mid 2000s as an indicator of success, the Nats have at least another three years of losing baseball before they will return to relevancy.

Graph 3 also allowed me to look at what 'blowing up' a roster does to the record of a team. In the Nats' case, hitting the reset button evidently has a direct negative impact on the team's record; each time the team partook in a sell-off, its record plummeted at a commensurate rate.

I thought at first I'd look at when a team decides to rebuild, but quickly realized that I couldn't due to not having the data I needed. I then changed my thinking to analyze the relationship between trading players away and record. What I struggled with, though, was what record to compare with which roster. Initially, I tried to look at the change opening day rosters between two seasons and the record of the most recent season, but this information was difficult to graph and understand, an idea highlighted in the code block below. 

#Player change since the Nats won the world series (2019)
```{r}
# OD Full Season Rosters
nats_od_roster_2018 <- mlb_rosters(team_id = 120, season = 2018, date = "2018-04-05", roster_type = 'fullSeason')
nats_od_roster_2019 <- mlb_rosters(team_id = 120, season = 2019, date = "2019-03-28", roster_type = 'fullSeason')
nats_od_roster_2020 <- mlb_rosters(team_id = 120, season = 2020, date = "2020-07-23", roster_type = 'fullSeason')
nats_od_roster_2021 <- mlb_rosters(team_id = 120, season = 2021, date = "2021-04-06", roster_type = 'fullSeason')
nats_od_roster_2022 <- mlb_rosters(team_id = 120, season = 2022, date = "2021-04-07", roster_type = 'fullSeason')

## Departures
nats_od_18_to_19 <- anti_join(nats_od_roster_2018, nats_od_roster_2019, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
nats_od_19_to_20 <- anti_join(nats_od_roster_2019, nats_od_roster_2020, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
nats_od_20_to_21 <- anti_join(nats_od_roster_2020, nats_od_roster_2021, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))
nats_od_21_to_22 <- anti_join(nats_od_roster_2021, nats_od_roster_2022, by = "person_full_name", copy = FALSE, na_matches = c("na", "never"))

## Counting & Pct Departures
nats_18_19_departures <- nats_od_18_to_19 %>% group_by(season) %>% summarize(departures = n()) %>% mutate(full_season = nats_od_roster_2018 %>% nrow()) %>% mutate(pct_departed = departures/full_season)
nats_19_20_departures <- nats_od_19_to_20 %>% group_by(season) %>% summarize(departures = n()) %>% mutate(full_season = nats_od_roster_2019 %>% nrow()) %>% mutate(pct_departed = departures/full_season)
nats_20_21_departures <- nats_od_20_to_21 %>% group_by(season) %>% summarize(departures = n()) %>% mutate(full_season = nats_od_roster_2020 %>% nrow()) %>% mutate(pct_departed = departures/full_season)
nats_21_22_departures <- nats_od_21_to_22 %>% group_by(season) %>% summarize(departures = n()) %>% mutate(full_season = nats_od_roster_2021 %>% nrow()) %>% mutate(pct_departed = departures/full_season)
nats_22_23_departures <- nats_od_21_to_22 %>% group_by(season) %>% summarize(departures = n()) %>% mutate(full_season = nats_od_roster_2021 %>% nrow()) %>% mutate(pct_departed = departures/full_season)

#Total players lost
nats_total_players_lost <- bind_rows(nats_19_20_departures, nats_20_21_departures, nats_21_22_departures) %>%
  rename("Year" = "season")

#Total players lost + win%
nats_players_lost_plus_win_percentage <- inner_join(nats_total_players_lost, nats_record_wins_losses_separate, by = "Year")

#Graphing total players lost + win%
ggplot() + geom_point(data=nats_players_lost_plus_win_percentage, aes(x=win_percentage, y=pct_departed))

#Records for 2018-2022 seasons
nats_2018_record <- bref_team_results('WSN', 2018) %>%
    filter(Gm == '162')
nats_2019_record <- bref_team_results('WSN', 2019) %>%
    filter(Gm == '162')
nats_2020_record <- bref_team_results('WSN', 2020) %>%
    filter(Gm == '60')
nats_2021_record <- bref_team_results('WSN', 2021) %>%
    filter(Gm == '162')
nats_2022_record <- bref_team_results('WSN', 2022) %>%
    filter(Gm == '162')

#Nats four-year record
nats_record_18_22 <- bind_rows(nats_2018_record, nats_2019_record, nats_2020_record, nats_2021_record, nats_2022_record)

#Separating the Record column into Wins and Losses
nats_record_wins_losses_separate <- separate(nats_record_18_22, col=Record, into=c('Wins', 'Losses'), sep='-') %>%
  mutate(Wins = as.numeric(Wins), Losses = as.numeric(Losses)) %>%
  mutate(win_percentage = (Wins/(Wins+Losses)*100)) %>% select(Year, Wins, win_percentage) %>% 
  mutate(previous_year_wins = case_when(
   Year == 2019 ~ 82,
   Year == 2020 ~ 93,
   Year == 2021 ~ 26,
   Year == 2022 ~ 65
   )
  ) %>% 
  mutate(win_change = previous_year_wins - Wins)

#Players lost since WS title
nats_players_lost <- bind_rows(nats_od_19_to_20, nats_od_20_to_21, nats_od_21_to_22)

#Graphing the wins by year using a line chart
ggplot(nats_record_wins_losses_separate, aes(x=Year, y=win_percentage)) +         
  geom_line() +
  geom_point()
```